---
title: "Assignment-1 Q&A Bilge Salman (03796071), Salim Kaplan (0378856)"
format: pdf
---

```{r}
# load packages here
library(tidyverse)
library(tidyr)
library(stats)

theme_set(theme_light())
```

# Task Set 1

For tasks 1.1-1.3, assume you throw 3 dice -- normal dice with 6 sides each ;)

## Task 1.1

Create a data frame with all possible combinations (outcomes) that can result from throwing all the dice. (Each row should represent one possible outcome.) Print the first and last 10 rows of the data frame and state how many possible outcomes there are.

```{r}
# df for all possible combinations of outcomes from throwing three dice
all_combinations <- expand.grid(die1 = 1:6, die2 = 1:6, die3 = 1:6)

# the first and last 10 rows of the df
print(head(all_combinations, 10))
print(tail(all_combinations, 10))

# Number of possible outcomes
num_outcomes <- nrow(all_combinations)
cat("Number of possible outcomes:", num_outcomes)
```

## Task 1.2

Create a data frame showing all possible sums that can result from throwing the three dice along with their probabilities. Report the results in a summary table (data frame) and a plot (visual graph).

```{r}
# Calculate the sum of each row
all_combinations$sum <- rowSums(all_combinations)

# Calculate the probabilities of each sum
sum_probabilities <- as.data.frame(table(all_combinations$sum))
colnames(sum_probabilities) <- c("Sum", "Frequency")
sum_probabilities$Probability <- sum_probabilities$Frequency / num_outcomes

# the summary table
sum_probabilities

# plot
ggplot(sum_probabilities, aes(x = Sum, y = Probability)) +
  geom_bar(stat = "identity", fill = 'blue') +
  labs(title = "Probability distribution of sums from throwing three dice",
       x = "Sum",
       y = "Probability")
```

## Task 1.3

Compute the probability that the sum is $\geq$ `10`, *given* that at least one of the dice shows a `3`.

```{r}
# filtering the outcomes where at least one die shows a 3
outcomes_with_3 <- subset(all_combinations, die1 == 3 | die2 == 3 | die3 == 3)

# calculating the total number of the outcomes
total_outcomes_with_3 <- nrow(outcomes_with_3)

# outcomes where the sum is greater than 9
outcomes_with_sum_ge_10 <- subset(outcomes_with_3, sum >= 10)
num_outcomes_with_sum_ge_10 <- nrow(outcomes_with_sum_ge_10)

# probability calculation
probability_sum_ge_10_given_3 <- num_outcomes_with_sum_ge_10 / total_outcomes_with_3
cat("Probability that the sum is >= 10 given that at least one die shows a 3 is:", round(probability_sum_ge_10_given_3, 2))
```

\newpage

# Task Set 2

For Task 2.1-2.3, assume you toss a globe `10` times, leading to either land or water.

## Task 2.1

Compute the probability of all possible numbers of occurrence of land, given the candidate proportion of `.5`. Report the results in a summary table and a plot and indicate whether the plot shows a probability distribution or a likelihood function.

```{r}
#write code here
```

## Task 2.2

Assume you observe `7` water. Take the candidate proportions of land `cp =` `r seq(0,1,.1)`. For each of these candidates, compute the probability of observing `7` water. Report the results in a summary table and a plot and indicate whether the plot shows a probability distribution or a likelihood function.

```{r}
#write code here
```

## Task 2.3

For each candidate proportion of land, compute the probability of all possible number of occurrences of land. Report the results in a summary table, showing the probability distributions as columns and the likelihood functions as rows.

```{r}
#write code here
```
